---
title: "WordleCode"
author: "Matt Ming"
date: "2/8/2022"
output: html_document
---

# Set Up

## Loading Dictionaries

First, create a function to read in the dictionary files and convert them to
single-column vectors

```{r}
load_dictionary <- function(filename){
  f <- read.csv(filename,header = TRUE)
  f <- f[,1] #Turn the one-column data frame from read.csv into a vector
  return(f)
}
```

Now, loading in both dictionaries as **valid_list** (Collin's Scrabble dictionary)
and **solution_list** (10000 common words dictionary)

```{r}
valid_list <- load_dictionary("collins-scrabble-words-2019.txt")
solution_list <- load_dictionary("google-10000-english-usa-no-swears.txt")
```

Narrowing down the **solution_list** to only contain words found in 
**valid_list**

```{r}
solution_list <- intersect(solution_list,valid_list)
```

After narrowing **solution_list**, it goes from containing 9884 words to 8336
words

## Getting solution word

Now, we want to create a function **pick_solution** which will get a 5-letter
word from the list of solution words and split that word into a vector of its 
characters

```{r}
pick_solution <- function(solutions,word_length = 5){
  #First filter all possible solution words to include only those which are
  #the correct length
  lengthlist <- solutions[which(nchar(solutions) == word_length)]
  #Choose one word from the list of correct-length words
  solword <- sample(lengthlist,1)
  #Split the word into a vector containing each of the letters in the word
  solvec <- strsplit(solword,c())[[1]]
  return(solvec)
}
solution <- pick_solution(solution_list,5) #Choose a solution word
#Note that we will also pick a new solution later when running the game
```

# Playing the Game

## Helper Function

We'll want to first construct a helper function **evaluate_guess** which
evaluates the user's guess and outputs a vector showing if each position of the 
guess was an exact match to the letter of the solution, was contained in the 
solution but not in the correct spot, or was not in the solution at all

```{r}
evaluate_guess <- function(guess,sol){
  #Initialize a vector which will store our evaluation characters
  eval_vec <- c()
  
  eval_vec[guess == sol] <- "*" #First check which characters in "guess" are
  #an exact match with the characters in "sol"
  eval_vec[!(guess %in% sol)] <- "-" #Check which characters in "guess" are
  #not in "sol"
  sol[sol == guess] <- "-" #Change characters matching in "sol" to "-"
  #so that they don't get double-counted
  
  #Set up a for loop to go through the last unchecked characters in "guess"
  for(i in which(is.na(eval_vec))){
    if(guess[i] %in% sol){ #If the remaining unchecked characters are in "sol"
      eval_vec[i] <- "+"
      sol[match(guess[i],sol)] <- "-" #Change character in sol to "-" so that
      #it doesn't get double counted
    }else{eval_vec[i] <- "-"} #If the remaining unchecked characters are not
    #in "sol"
  }
  return(eval_vec)
}
```

## Main Game

We'll now start setting up the code for actually playing the game.  The function
**play_wordle** should take arguments *solution*, *valid_list*, *num_guesses*, 
and *word_length*, and do the following:

1. Tell the players the rules: "You have [*nun_guesses*] chances to guess a word 
of length [*word_length*]"

2. Display the letters not yet guessed as "Possible Letters:..."

3. Prompt the player for a guess and check if the guess is valid (i.e., correct
length, no numbers/non-letter characters, is a valid word), and convert the 
guess to a vector of letters (same as the solution vector)

4. Compare the guess w/ the solution using the **evaluate_guess** helper 
fucntion created above

5. Update the list of not-yet-guessed letters

6. Check if the puzzle was solved.  If the player got the word correct, give a
"YOU WON" message as well as guess and feedback history.  If the player was
not correct, prompt for another guess

7. If the player has gone past the number of allotted guesses and is still not
correct, return a "YOU LOSE" message and give the guess and feedback history

```{r}
play_wordle <- function(solution,valid_list,num_guesses = 6,word_length = 5){
  #Initializing function
  guesses <- 1 #Initialize a variable for counting number of guesses
  letters_left <- LETTERS #Initialize a vector containing all letters yet to be
  #guessed (starts out containing all capital English letters)
  guess_hist <- list() #Initialize list which will store all user guesses
  eval_hist <- list() #Initialize list which will store all evaluations
  
  #Formatting for making the game presentation look nice
  cat("\n\nWelcome to Matt's Wordle!\n\n")
  #Explaining the rules of Wordle
  cat("You have",num_guesses,"chances to guess a word of length",word_length)
  cat("\nSymbols Legend:") #Explaining the evaluation symbols
  cat("\n* = Letter is in correct spot")
  cat("\n+ = Letter is in solution but in a different spot")
  cat("\n- = Letter is not in solution")
  cat("\n\n\n")
  
  #Set up loop to go until the game is declared over (win or lose)
  gameover <- FALSE #Set up another variable to keep track of if the game is
  #over (i.e., the player has won or lost)
  while(gameover == FALSE){
    #Formatting
    cat("Guess #",guesses,sep = "") #Display current guess number
    cat("\nPossible Letters:",letters_left) #Display list of possible letters
    cat("\n")
    
    #Get user input for a guess
    guess <- readline(paste0("Enter a ",word_length,
                             "-letter word here, then press <ENTER>: "))
    
    #Set up a check for if the user's guess is a valid word
    is_valid <- FALSE
    #Continue to ask for user input as long as user supplies an invalid guess
    #Entering an invalid word should not count towards your guess number
    while(is_valid == FALSE){
      if(nchar(guess) > word_length){ #Guess was too long
        guess <- readline(paste0("Your guess was too long. Enter a ",
                                 word_length,
                                 "-letter word here, then press <ENTER>: "))
      }else if(nchar(guess) < word_length){ #Guess was too short
        guess <- readline(paste0("Your guess was too short. Enter a ",
                                 word_length,
                                 "-letter word here, then press <ENTER>: "))
      }else if(!(toupper(guess) %in% valid_list)){ #Guess was not a valid word
        guess <- readline(paste0("Your guess was not a valid word. Enter a ",
                                 word_length,
                                 "-letter word here, then press <ENTER>: "))
      }else{is_valid <- TRUE} #If no invalid errors, continue to next step
    }
    
    #Convert guess into vector of characters
    guess <- strsplit(toupper(guess),c())[[1]]
    guess_hist[[guesses]] <- guess #Add guess to list of guesses
    cat("Your Guess:",guess)
    eval <- evaluate_guess(guess,solution) #Evaluate the guess
    eval_hist[[guesses]] <- eval #Add evaluation to list of evaluations
    cat("\nEvaluation:",eval)
    
    #Checks for the end of the game
    if(length(which(eval == "*")) == word_length){ #If all evaluations are "*"
      #meaning the user correctly guessed the word
      cat("\n\n\n")
      cat("YOU WON!")
      cat("\nYou used",guesses,"guesses") #Show number of guesses needed
      gameover <- TRUE #Exit while loop
    }else if(guesses < num_guesses){ #If guess is not correct but user still
      #has guesses remaining
      guesses <- guesses + 1 #Increase guess count by 1
      #Change list of possible letters to remove the ones which were guessed and
      #not part of the solution
      letters_left <- setdiff(letters_left,guess[which(eval == "-")])
      letters_left <- union(letters_left,
                            guess[which(eval == "+" | eval == "*")])
      letters_left <- sort(letters_left)
      cat("\n\n")
    }else{ #If user is not correct and is out of guesses
      cat("\n\n\n")
      cat("YOU LOSE :(")
      cat("\nThe correct word was ")
      cat(solution,sep = "")
      gameover <- TRUE #Exit while loop
    }
  }
  #Formatting
  cat("\n\n\n")
  #Print a list of all of user's guesses and evaluations
  cat("Your guesses:")
  cat("\n")
  for(i in 1:length(guess_hist)){
    cat("\nGuess #",i,": ",sep = "")
    cat(guess_hist[[i]])
    cat("\n          ",sep = "")
    cat(eval_hist[[i]])
  }
}
```

# Play Time!

Now, actually running a play of the game -- running the following chunk of code
will allow you to actually play (Note that the arguments which take numbers have
default values, so in this case the numbers are optional):

```{r,eval=FALSE}
solution <- pick_solution(solution_list,5) #Choose a solution word
play_wordle(solution,valid_list,6,5) #Start Wordle game
```

Note that because of the use of user input in the **play_wordle** function, the
above code is not evaluated so that the .Rmd knit is able to work

Have fun!